---
title: "Converting rice microarray probe IDs with OryzaProbe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Converting rice microarray probe IDs with OryzaProbe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

OryzaProbe converts expression-matrix row names from three rice microarray
platforms: GPL6864, GPL2025, and GPL8852. Version 0.2.0 preserves the input row
order and records how each probe was resolved.

```{r}
library(OryzaProbe)
supported_platforms()
```

## Convert and merge probes

The two GPL6864 probes below map to the same RAP-DB gene. Supplying the platform
explicitly is recommended for reproducible analysis.

```{r}
expression <- matrix(
  c(2, 4, 6, 8),
  nrow = 2,
  dimnames = list(c("36144", "37310"), c("sample_1", "sample_2"))
)

converted <- probe_convert(
  expression,
  platform = "GPL6864",
  merge_by = "mean"
)
converted
```

The result is a numeric matrix. The report links every input probe to its
annotation, resolution policy, and output row.

```{r}
probe_conversion_report(converted)
probe_conversion_summary(converted)
```

## Automatic platform detection

Automatic detection scores every input row. It therefore works for valid
subsets that do not contain one particular sentinel probe.

```{r}
subset_expression <- matrix(
  c(13, 14),
  ncol = 1,
  dimnames = list(c("13", "14"), "sample")
)
probe_convert(subset_expression)
```

Detection raises an error if fewer than half of the probe IDs match a supported
platform or if the result is ambiguous.

## Unmapped, non-RAP, and multi-target annotations

Platform annotations contain controls, miRNA identifiers, probes without a RAP
target, and a few one-to-many mappings. These are not silently discarded.

The `unmapped` and `non_rap` arguments accept `"keep"`, `"drop"`, or `"error"`.
The `multi_mapping` argument accepts `"keep"`, `"first"`, `"expand"`, `"drop"`,
or `"error"`.

```{r}
multi <- matrix(
  10,
  dimnames = list("Os.19863.1.S1_a_at", "sample")
)
probe_convert(
  multi,
  platform = "GPL2025",
  multi_mapping = "expand"
)
```

## Annotation provenance

The GPL6864 and GPL2025 mappings are the tables published with OryzaProbe 0.1.0.
They remain identifiable as a legacy snapshot rather than being presented as a
current RAP-DB release.

```{r}
probe_annotation_metadata()
head(probe_annotation("GPL6864"))
```

The original `probeConvert()` function remains available for existing scripts,
but new work should use `probe_convert()` so mapping policies are visible.
